Experimentally informed site-specific substitution models deepen phylogenetic estimates of the divergence of viral lineages

### 2018-02-12 manuscript/manuscript.tex
first draft of competing

\skhcomment{Section Outline:}

\begin{itemize}
\item We investigated the competing effects of ``shifting preferences" and long branches by comparing ExpCMs with trees of increasing overall divergence from the focal sequence. 
\item We use the fit stringency parameter $\beta$ as a measure of relevance. 
\begin{itemize}
\item The stringency parameter rescales the preferences. A stringency parameter with a value greater than one indicates that selection in nature prefers the same amino acids as selection in lab but with greater strength. 
\item There is an inverse relationship between the stringency parameter and overall sequence divergence. As the overall divergence of the tree increases, the stringency parameter decreases. 
\end{itemize}
\item \skhcomment{hypothetical data} There is also a positive relationship between the stringency parameter and branch lengths 
\begin{itemize}
\item We optimized the same tree with ExpCMs with different stringency parameters. 
\item The ExpCMs with higher stringency values estimate longer branches than the ExpCMs with lower stringency values. 
\end{itemize}
\end{itemize}

@@

\skhcomment{Section Outline}

\begin{itemize}
\item \texttt{phylobayes} implements the mutation-selection model in a bayesian framework as a way to avoid overfitting. 
\item \texttt{phylobayes} and ExpCM (H1+H3 average) estimate branches of roughly equal length 
\begin{itemize}
\item the ExpCM estimate longer branches from the focal sequences 
\item \texttt{phylobayes} estimates longer branches from the non-focal sequences 
\end{itemize}
\end{itemize}

>>>>

We further investigated the effect of ``shifting" preferences by comparing behavior of ExpCM across trees with increasing overall divergence from the focal sequence of the deep mutational scans. 
Based on our results with the large HA trees, we expected the preferences to be more "relevant" on trees which have a low overall sequence divergence from the DMS focal sequence.

We use the value of the ExpCM stringency parameter to measure the ``relevance" of a given DMS preference set and tree. 
A stringency parameter with a value greater than one indicates that selection in nature prefers the same amino acids as selection in lab but with greater strength, or that the preferences are ``relevant" to evolution in nature. 
Conversely, when the stringency parameter is fit to a value close to zero, the preferences are not ``relevant" and de-emphasize the measurements from the lab so that the preferences are flat. 
Using the ExpCM (H1 preferences), we observe an inverse relationship between overall divergence and the stringency. 
As the tree includes more and more sequences which are further from the focal sequence, the relevance of the preferences declines. 
This indicates that the preferences truly are most relevant on trees with sequences closely related to the focal sequence (short branches).

While the preferences are most relevant for the trees with short branches, due to the low divergence from the focal sequence, the effect of site-specific stationary states on branch length estimation is most important on long branches. 
On the low divergence tree, with the most ``relevant" preferences, roughly the same branch length estimated by the GY94+$\Gamma\omega$  and the ExpCM+$\Gamma\omega$. 
In fact, it is not until the high divergence tree that a true effect is seen, in the branches from the focal sequences. 
The effect from site-specific substitution models doesn't really have an effect until the branch lengths are a particular length.

This leads to a regime with competing effects. 
As the branch lengths increase, the ``importance" of having a site-specific stationary state increases. 
At the same time, the ``relevance" of an given site-specific stationary decreases as the divergence increases. 
Current models which have one stationary state across the entire tree, site-specific or not, will not be able to address this tension.

@@

Finally, we compared the branch lengths estimated by ExpCM to branch lengths estimated by another site-specific stationary state model. 
We inferred the branch lengths on the ``high" divergence tree using the mutation-selection model implement in \texttt{phylobayes}. 
Unlike the ExpCM, this mutation-selection model in \texttt{phylobayes}
